Untitled

### 2022-06-07 Manuscript/Lanuza_et_al_2021.Rmd
edit

The phylogenetically informed principal component analysis captured by the first two and three axes 51.8% and 70.97% of the reproductive trait variation, respectively (Fig. 1 and Fig. S7). The first principal component (PC1) represented 26.72% of the trait variation and indicated a negative correlation between flower number and flower size ('flower number - flower size'). The main contributing traits to PC1 were plant height, flower number, ovule number and flower size (loadings \> \|0.5\|; Table S3) but style length also contributed moderately to PC1 (loading = -0.33). One end of this axis comprised species with high investment in flower number and plant height but small flower size, short style length and low ovule number. For instance, on this end of the spectrum we find the species *Cornus florida* which has approximately a total of 10.000 flowers, an average height of 7.5m, flowers of 3mm wide, a style length of 3.5mm and a total of 2 ovules per flower. The other end of this spectrum comprised species that were short in height and invested in large flowers, long styles, many ovules, but few flowers. For instance, on this side of the axis we find the species *Petunia axillaris* that has approximately 10 flowers per plant, a height of 0.5m, flowers over 50mm wide, styles of 25mm and over 200 ovules per flower. The second principal component (PC2) represented 25.05% of the trait variation and indicated the variation from low to high autonomous selfing, or in other words, high to low pollinator dependence, respectively. The main driver of trait variation on PC2 was autonomous selfing (loading = 0.85) but the other traits (except ovule number) also made moderate contributions (loadings from 0.27 to 0.4; Table S3). In general terms, species with high pollinator dependence were associated with larger and a higher number of flowers, greater plant height and longer styles. A species example of this side of the spectrum is *Zuccagnia punctata* that is a self incompatible shrub which depends totally on floral visitors, it has approximately 1500 flowers per plant, 3m height and a style length of 20mm. In contrast, species with low pollinator dependence tended to have fewer and smaller flowers, shorter plant height and shorter styles. For instance, on this side of the spectrum we find the species *Veronica peregrina* which is a self compatible herb that is thought to be almost a complete selfer, it has around 20 flowers per plant, a height of 0.2m and a style length of 0.25mm. Further, PC3 explained a considerable amount of trait variability (19.17%) and the main contributors to this axis were style length (loading = -0.66) and the degree of autonomous selfing (loading = -0.51). The remaining traits, apart from ovule number, were moderately correlated to changes on PC3 (loadings from -0.23 to -0.46; Table S3). In addition, the pPCA with the subset of species that had nectar and pollen quantity data showed that nectar quantity (microlitres of nectar per flower) and pollen grains per flower were positively associated with flower size, style length and ovule number but negatively associated with flower number (PC1, 26.82%; Fig. S4). This pPCA explained similar variance with the first two principal components (45.52%) and similar associations of traits despite some variability in the loadings (Table S4).

We found that most categorical traits were statistically associated with the first two axes of trait variation (Fig. 2 and Table S2). Flower symmetry, which was only associated with PC2 (Sum of squares = 8.51, F-value = 14.72, *P* \< 0.01 ), and nectar provision, which was independent of PC1 and PC2 (PC1: Sum of squares = 0.37, F-value = 0.29 , *P* = 0.59; PC2: Sum of squares = 0.83, F-value = 1.43, *P* = 0.23) showed lack of statistical association. In addition, we found (with a Tukey test) statistical differences between the different levels of categorical traits in the trait space (Fig. S8). Regarding self compatibility, we found larger differences on PC2 (i.e., species with unisexual flowers that were self incompatible were statistically differentiated from species with partial or full self compatibility; Fig. S8a and Fig. S8b). Life forms differed statistically across both axes of trait variation and followed a gradient of larger life forms (trees and shrubs) with higher pollinator dependence to smaller ones (herbs) with lower pollinator dependence (Fig. S8c and Fig. S8d). Consequently, lifespan also followed this gradient but perennial and short lived species only differed statistically on PC2 (Fig. S8e and Fig. S8f). Species with unisexual flowers (monoecious and dioecious) were clustered on both extremes of the first two principal components and had the highest pollinator dependence and highest number of flowers (Fig. S8g and Fig. S8h). Moreover, we found that the campanulate and capitulum flower shapes were differentiated from tube, papilionaceous, open and brush shapes in the trait space. The former morphologies had larger flowers and greater pollinator dependence, while the latter had higher flower number and greater autonomous selfing (Fig. S8i and Fig. S8j). Regarding flower symmetry, zygomorphic flowers were associated with lower levels of pollinator dependence, whereas actinomorphic flowers had higher levels of pollinator dependence (Fig. S8k and Fig. S8l).

@@

The main axes of trait variation explained partly presence-absence of interaction partners (conditional $R^{2}$ = 0.26; marginal $R^{2}$ = 0.20) but little of the overall number of visits (conditional $R^{2}$ = 0.31; marginal $R^{2}$ = 0.06). However, we found relevant differences across the different floral visitor guilds on both presence-absence of interactions and number of visits (Fig. 3). We found that plants with high flower number and small flowers had higher interaction partners of Coleoptera, non-bee-Hymenoptera and all Diptera guilds but plants with low flower number but large flowers had higher interaction partners of bees and Lepidoptera guilds (flower number - flower size trade-off, PC1; Fig. 3a). However, plant species with high flower number but small flowers had higher number of visits of bees and syrphids guilds (PC1; Fig. 3d). Remarkably, all plant species with higher pollinator dependence had higher number of interacting partners and number of visits for all floral visitor guilds (PC2; Fig. 3b and Fig. 3e). Finally, plant species with short styles and low selfing had higher interaction partners of all guilds but bees that interacted clearly more with plant species with long styles and high selfing (style length trade-off; Fig. 3c). However, for number of visits, we found that plants with long styles and high selfing interacted more frequently with Lepidoptera and non-bee-Hymenoptera guilds (Fig. 3f).

>>>>

The phylogenetically informed principal component analysis captured by the first two and three axes 51.8% and 70.97% of the reproductive trait variation, respectively (Fig. 1 and Fig. S7). The first principal component (PC1 or the flower number - flower size axis) represented 26.72% of the trait variation and indicated a negative correlation between flower number and flower size. The main contributing traits to PC1 were plant height, flower number, ovule number and flower size (loadings \> \|0.5\|; Table S3) but style length also contributed moderately to PC1 (loading = -0.33). One end of this axis comprised species with high investment in flower number and plant height but small flower size, short style length and low ovule number. For instance, on this end of the spectrum we find the species *Cornus florida* which has approximately a total of 10.000 flowers, an average height of 7.5m, flowers of 3mm wide, a style length of 3.5mm and a total of 2 ovules per flower. The other end of this spectrum comprised species that were short in height and invested in large flowers, long styles, many ovules, but few flowers. For instance, on this side of the axis we find the species *Petunia axillaris* that has approximately 10 flowers per plant, a height of 0.5m, flowers over 50mm wide, styles of 25mm and over 200 ovules per flower. The second principal component (PC2 or the pollinator dependence axis) represented 25.05% of the trait variation and indicated the variation from low to high autonomous selfing, or in other words, high to low pollinator dependence, respectively. The main driver of trait variation on PC2 was autonomous selfing (loading = 0.85) but the other traits (except ovule number) also made moderate contributions (loadings from 0.27 to 0.4; Table S3). In general terms, species with high pollinator dependence were associated with larger and a higher number of flowers, greater plant height and longer styles. On this extreme of the spectrum we find the species *Zuccagnia punctata* that is a self-incompatible shrub that depends completely on floral visitors for seed production, it has approximately 1500 flowers per plant, 3m height and a style length of 20mm. In contrast, species with low pollinator dependence tended to have fewer and smaller flowers, shorter plant height and shorter styles. For instance, on this end of the spectrum we find the species *Veronica peregrina* which is a self compatible herb that is thought to be almost a complete selfer that relies none or little on floral visitors, it has around 20 flowers per plant, a height of 0.2m and a style length of 0.25mm. Further, the third principal component (PC3 or the style length - pollinator dependence axis) explained a considerable amount of trait variability (19.17%) and represented a negative correlation between style length (loading = -0.66) and pollinator dependence (autonomous selfing loading = |-0.51|) where species with short styles had high pollinator dependence and species with long styles low pollinator dependence. The remaining traits, apart from ovule number, were positively correlated with style length and negatively correlated with pollinator dependence (loadings from -0.23 to -0.46; Table S3). In addition, the pPCA with the subset of species that had nectar and pollen quantity data showed that nectar quantity (microlitres of nectar per flower) and pollen grains per flower were positively associated with flower size, style length and ovule number but negatively associated with flower number (PC1, 26.82%; Fig. S4). This pPCA explained similar variance with the first two principal components (45.52%) and similar associations of traits despite some variability in the loadings (Table S4).

We found (with an Anova) that most categorical traits were statistically associated with the first two axes of trait variation (Fig. 2 and Table S2). Flower symmetry, which was only associated with PC2 (Sum of squares = 8.51, F-value = 14.72, *P* \< 0.01 ), and nectar provision, which was independent of PC1 and PC2 (PC1: Sum of squares = 0.37, F-value = 0.29 , *P* = 0.59; PC2: Sum of squares = 0.83, F-value = 1.43, *P* = 0.23) showed lack of statistical association. In addition, we found (with a Tukey test) statistical differences between the different levels of categorical traits in the trait space (Fig. S8). Regarding self compatibility, we found larger differences on PC2 (i.e., species with unisexual flowers that were self incompatible were statistically differentiated from species with partial or full self compatibility; Fig. S8a and Fig. S8b). Life forms differed statistically across both axes of trait variation and followed a gradient of larger life forms (trees and shrubs) with higher pollinator dependence to smaller ones (herbs) with lower pollinator dependence (Fig. S8c and Fig. S8d). Consequently, lifespan also followed this gradient but perennial and short lived species only differed statistically on PC2 (Fig. S8e and Fig. S8f). Species with unisexual flowers (monoecious and dioecious) were clustered on both extremes of the first two principal components and had the highest pollinator dependence and highest number of flowers (Fig. S8g and Fig. S8h). Moreover, we found that the campanulate and capitulum flower shapes were differentiated from tube, papilionaceous, open and brush shapes in the trait space. The former morphologies had larger flowers and greater pollinator dependence, while the latter had higher flower number and greater autonomous selfing (Fig. S8i and Fig. S8j). Regarding flower symmetry, zygomorphic flowers were associated with lower levels of pollinator dependence, whereas actinomorphic flowers had higher levels of pollinator dependence (Fig. S8k and Fig. S8l).

@@

The main axes of trait variation explained partly presence-absence of interaction partners (conditional $R^{2}$ = 0.26; marginal $R^{2}$ = 0.20) but little of the overall number of visits (conditional $R^{2}$ = 0.31; marginal $R^{2}$ = 0.06). However, we found relevant differences across the different floral visitor guilds on both presence-absence of interactions and number of visits (Fig. 3). We found on the flower number - flower size axis that plants with high flower number and small flowers had higher interaction partners of Coleoptera, non-bee-Hymenoptera and all Diptera guilds (Fig. 3a) but plants on the other extreme of the spectrum with low flower number but large flowers had higher interaction partners of bees and Lepidoptera guilds. Regarding visitation rates, we found similar trends but bees and syrphids guilds in this case showed higher number of visits to plant species with high flower number and small flowers (Fig. 3d). On the pollinator dependence axis, we found that all species with higher pollinator dependence had higher number of interacting partners and number of visits from all floral visitor guilds (Fig. 3b and Fig. 3e). Finally, on the style length - pollinator dependence axis plant species with short styles and high pollinator dependence had higher interaction partners of all guilds but bees, that interacted clearly more with plant species with long styles and low pollinator dependence (Fig. 3c). However, regarding number of visits, plants with long styles and high selfing interacted more frequently with Lepidoptera and non-bee-Hymenoptera guilds (Fig. 3f).

### 2022-06-07 Manuscript/Lanuza_et_al_2021.tex
edit

The phylogenetically informed principal component analysis captured by the first two and three axes 51.8\% and 70.97\% of the reproductive trait variation, respectively (Fig. 1 and Fig. S7). The first principal component (PC1) represented 26.72\% of the trait variation and indicated a negative correlation between flower number and flower size (`flower number - flower size'). The main contributing traits to PC1 were plant height, flower number, ovule number and flower size (loadings \textgreater{} \textbar0.5\textbar; Table S3) but style length also contributed moderately to PC1 (loading = -0.33). One end of this axis comprised species with high investment in flower number and plant height but small flower size, short style length and low ovule number. For instance, on this end of the spectrum we find the species \emph{Cornus florida} which has approximately a total of 10.000 flowers, an average height of 7.5m, flowers of 3mm wide, a style length of 3.5mm and a total of 2 ovules per flower. The other end of this spectrum comprised species that were short in height and invested in large flowers, long styles, many ovules, but few flowers. For instance, on this side of the axis we find the species \emph{Petunia axillaris} that has approximately 10 flowers per plant, a height of 0.5m, flowers over 50mm wide, styles of 25mm and over 200 ovules per flower. The second principal component (PC2) represented 25.05\% of the trait variation and indicated the variation from low to high autonomous selfing, or in other words, high to low pollinator dependence, respectively. The main driver of trait variation on PC2 was autonomous selfing (loading = 0.85) but the other traits (except ovule number) also made moderate contributions (loadings from 0.27 to 0.4; Table S3). In general terms, species with high pollinator dependence were associated with larger and a higher number of flowers, greater plant height and longer styles. A species example of this side of the spectrum is \emph{Zuccagnia punctata} that is a self incompatible shrub which depends totally on floral visitors, it has approximately 1500 flowers per plant, 3m height and a style length of 20mm. In contrast, species with low pollinator dependence tended to have fewer and smaller flowers, shorter plant height and shorter styles. For instance, on this side of the spectrum we find the species \emph{Veronica peregrina} which is a self compatible herb that is thought to be almost a complete selfer, it has around 20 flowers per plant, a height of 0.2m and a style length of 0.25mm. Further, PC3 explained a considerable amount of trait variability (19.17\%) and the main contributors to this axis were style length (loading = -0.66) and the degree of autonomous selfing (loading = -0.51). The remaining traits, apart from ovule number, were moderately correlated to changes on PC3 (loadings from -0.23 to -0.46; Table S3). In addition, the pPCA with the subset of species that had nectar and pollen quantity data showed that nectar quantity (microlitres of nectar per flower) and pollen grains per flower were positively associated with flower size, style length and ovule number but negatively associated with flower number (PC1, 26.82\%; Fig. S4). This pPCA explained similar variance with the first two principal components (45.52\%) and similar associations of traits despite some variability in the loadings (Table S4).

We found that most categorical traits were statistically associated with the first two axes of trait variation (Fig. 2 and Table S2). Flower symmetry, which was only associated with PC2 (Sum of squares = 8.51, F-value = 14.72, \emph{P} \textless{} 0.01 ), and nectar provision, which was independent of PC1 and PC2 (PC1: Sum of squares = 0.37, F-value = 0.29 , \emph{P} = 0.59; PC2: Sum of squares = 0.83, F-value = 1.43, \emph{P} = 0.23) showed lack of statistical association. In addition, we found (with a Tukey test) statistical differences between the different levels of categorical traits in the trait space (Fig. S8). Regarding self compatibility, we found larger differences on PC2 (i.e., species with unisexual flowers that were self incompatible were statistically differentiated from species with partial or full self compatibility; Fig. S8a and Fig. S8b). Life forms differed statistically across both axes of trait variation and followed a gradient of larger life forms (trees and shrubs) with higher pollinator dependence to smaller ones (herbs) with lower pollinator dependence (Fig. S8c and Fig. S8d). Consequently, lifespan also followed this gradient but perennial and short lived species only differed statistically on PC2 (Fig. S8e and Fig. S8f). Species with unisexual flowers (monoecious and dioecious) were clustered on both extremes of the first two principal components and had the highest pollinator dependence and highest number of flowers (Fig. S8g and Fig. S8h). Moreover, we found that the campanulate and capitulum flower shapes were differentiated from tube, papilionaceous, open and brush shapes in the trait space. The former morphologies had larger flowers and greater pollinator dependence, while the latter had higher flower number and greater autonomous selfing (Fig. S8i and Fig. S8j). Regarding flower symmetry, zygomorphic flowers were associated with lower levels of pollinator dependence, whereas actinomorphic flowers had higher levels of pollinator dependence (Fig. S8k and Fig. S8l).

@@

The main axes of trait variation explained partly presence-absence of interaction partners (conditional \(R^{2}\) = 0.26; marginal \(R^{2}\) = 0.20) but little of the overall number of visits (conditional \(R^{2}\) = 0.31; marginal \(R^{2}\) = 0.06). However, we found relevant differences across the different floral visitor guilds on both presence-absence of interactions and number of visits (Fig. 3). We found that plants with high flower number and small flowers had higher interaction partners of Coleoptera, non-bee-Hymenoptera and all Diptera guilds but plants with low flower number but large flowers had higher interaction partners of bees and Lepidoptera guilds (flower number - flower size trade-off, PC1; Fig. 3a). However, plant species with high flower number but small flowers had higher number of visits of bees and syrphids guilds (PC1; Fig. 3d). Remarkably, all plant species with higher pollinator dependence had higher number of interacting partners and number of visits for all floral visitor guilds (PC2; Fig. 3b and Fig. 3e). Finally, plant species with short styles and low selfing had higher interaction partners of all guilds but bees that interacted clearly more with plant species with long styles and high selfing (style length trade-off; Fig. 3c). However, for number of visits, we found that plants with long styles and high selfing interacted more frequently with Lepidoptera and non-bee-Hymenoptera guilds (Fig. 3f).

>>>>

The phylogenetically informed principal component analysis captured by the first two and three axes 51.8\% and 70.97\% of the reproductive trait variation, respectively (Fig. 1 and Fig. S7). The first principal component (PC1 or the flower number - flower size axis) represented 26.72\% of the trait variation and indicated a negative correlation between flower number and flower size. The main contributing traits to PC1 were plant height, flower number, ovule number and flower size (loadings \textgreater{} \textbar0.5\textbar; Table S3) but style length also contributed moderately to PC1 (loading = -0.33). One end of this axis comprised species with high investment in flower number and plant height but small flower size, short style length and low ovule number. For instance, on this end of the spectrum we find the species \emph{Cornus florida} which has approximately a total of 10.000 flowers, an average height of 7.5m, flowers of 3mm wide, a style length of 3.5mm and a total of 2 ovules per flower. The other end of this spectrum comprised species that were short in height and invested in large flowers, long styles, many ovules, but few flowers. For instance, on this side of the axis we find the species \emph{Petunia axillaris} that has approximately 10 flowers per plant, a height of 0.5m, flowers over 50mm wide, styles of 25mm and over 200 ovules per flower. The second principal component (PC2 or the pollinator dependence axis) represented 25.05\% of the trait variation and indicated the variation from low to high autonomous selfing, or in other words, high to low pollinator dependence, respectively. The main driver of trait variation on PC2 was autonomous selfing (loading = 0.85) but the other traits (except ovule number) also made moderate contributions (loadings from 0.27 to 0.4; Table S3). In general terms, species with high pollinator dependence were associated with larger and a higher number of flowers, greater plant height and longer styles. On this extreme of the spectrum we find the species \emph{Zuccagnia punctata} that is a self-incompatible shrub that depends completely on floral visitors for seed production, it has approximately 1500 flowers per plant, 3m height and a style length of 20mm. In contrast, species with low pollinator dependence tended to have fewer and smaller flowers, shorter plant height and shorter styles. For instance, on this end of the spectrum we find the species \emph{Veronica peregrina} which is a self compatible herb that is thought to be almost a complete selfer that relies none or little on floral visitors, it has around 20 flowers per plant, a height of 0.2m and a style length of 0.25mm. Further, the third principal component (PC3 or the style length - pollinator dependence axis) explained a considerable amount of trait variability (19.17\%) and represented a negative correlation between style length (loading = -0.66) and pollinator dependence (autonomous selfing loading = \textbar-0.51\textbar) where species with short styles had high pollinator dependence and species with long styles low pollinator dependence. The remaining traits, apart from ovule number, were positively correlated with style length and negatively correlated with pollinator dependence (loadings from -0.23 to -0.46; Table S3). In addition, the pPCA with the subset of species that had nectar and pollen quantity data showed that nectar quantity (microlitres of nectar per flower) and pollen grains per flower were positively associated with flower size, style length and ovule number but negatively associated with flower number (PC1, 26.82\%; Fig. S4). This pPCA explained similar variance with the first two principal components (45.52\%) and similar associations of traits despite some variability in the loadings (Table S4).

We found (with an Anova) that most categorical traits were statistically associated with the first two axes of trait variation (Fig. 2 and Table S2). Flower symmetry, which was only associated with PC2 (Sum of squares = 8.51, F-value = 14.72, \emph{P} \textless{} 0.01 ), and nectar provision, which was independent of PC1 and PC2 (PC1: Sum of squares = 0.37, F-value = 0.29 , \emph{P} = 0.59; PC2: Sum of squares = 0.83, F-value = 1.43, \emph{P} = 0.23) showed lack of statistical association. In addition, we found (with a Tukey test) statistical differences between the different levels of categorical traits in the trait space (Fig. S8). Regarding self compatibility, we found larger differences on PC2 (i.e., species with unisexual flowers that were self incompatible were statistically differentiated from species with partial or full self compatibility; Fig. S8a and Fig. S8b). Life forms differed statistically across both axes of trait variation and followed a gradient of larger life forms (trees and shrubs) with higher pollinator dependence to smaller ones (herbs) with lower pollinator dependence (Fig. S8c and Fig. S8d). Consequently, lifespan also followed this gradient but perennial and short lived species only differed statistically on PC2 (Fig. S8e and Fig. S8f). Species with unisexual flowers (monoecious and dioecious) were clustered on both extremes of the first two principal components and had the highest pollinator dependence and highest number of flowers (Fig. S8g and Fig. S8h). Moreover, we found that the campanulate and capitulum flower shapes were differentiated from tube, papilionaceous, open and brush shapes in the trait space. The former morphologies had larger flowers and greater pollinator dependence, while the latter had higher flower number and greater autonomous selfing (Fig. S8i and Fig. S8j). Regarding flower symmetry, zygomorphic flowers were associated with lower levels of pollinator dependence, whereas actinomorphic flowers had higher levels of pollinator dependence (Fig. S8k and Fig. S8l).

@@

The main axes of trait variation explained partly presence-absence of interaction partners (conditional \(R^{2}\) = 0.26; marginal \(R^{2}\) = 0.20) but little of the overall number of visits (conditional \(R^{2}\) = 0.31; marginal \(R^{2}\) = 0.06). However, we found relevant differences across the different floral visitor guilds on both presence-absence of interactions and number of visits (Fig. 3). We found that plants with high flower number and small flowers had higher interaction partners of Coleoptera, non-bee-Hymenoptera and all Diptera guilds but plants with low flower number but large flowers had higher interaction partners of bees and Lepidoptera guilds (flower number - flower size trade-off, PC1; Fig. 3a). However, plant species with high flower number but small flowers had higher number of visits of bees and syrphids guilds (PC1; Fig. 3d). All plant species with higher pollinator dependence had higher number of interacting partners and number of visits for all floral visitor guilds (PC2; Fig. 3b and Fig. 3e). Finally, plant species with short styles and high pollinator dependence had higher interaction partners of all guilds but bees that interacted clearly more with plant species with long styles and low pollinator dependence (style length trade-off; Fig. 3c). However, for number of visits, we found that plants with long styles and high selfing interacted more frequently with Lepidoptera and non-bee-Hymenoptera guilds (Fig. 3f).

### 2022-06-07 Supplementary_Material/Sup_Mat_Lanuza_et_al_2021_Ecology_Letters.Rmd
edit

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=10, fig.width=10, fig.cap="\\textbf{Fig. S7.} Phylogenetic informed principal component analysis for the subset of species with floral rewards."}
setwd("~/R_Projects/Reproductive Traits")

#Load libraries
#LOAD LIBRARIES
library(phytools)
library(ape) #for phylogenetic distance
library(dplyr) #data processing
library(rtrees) #for phylogenetic distancelibrary(MASS)
library(reshape2)
library(viridis) #COLOUR GGPLOT
library(MASS)
library(ggplot2)
library(broman) #crayon colours
library(cowplot)
# Theme for publication
theme_ms <- function(base_size=12, base_family="Helvetica") {
  (theme_bw(base_size = base_size, base_family = base_family)+
      theme(text=element_text(color="black"),
            axis.title=element_text( size = rel(1.2)),
            axis.text=element_text(size = rel(1.75), color = "black"),
            legend.title=element_text(face="bold"),
            legend.text=element_text(),
            legend.background=element_rect(fill="transparent"),
            legend.key.size = unit(1.2, 'lines'),
            panel.border=element_rect(color="black",size=1),
            panel.grid.minor.x =element_blank(),
            panel.grid.minor.y= element_blank(),
             panel.grid.major= element_blank()
      ))
}

@@

@@

  #ADD LABELS
  rownames(PC$L) <- c("Selfing", "Flower number", "Flower Size", "Style length", "Ovule number", "Plant height", "Nectar","Pollen" )

  PCAloadings <- data.frame(Variables = rownames(PC$L), PC$L)
  plot <- plot + annotate("text", x = -(PCAloadings$PC1*c(4.85,4.8,4.8,5.5,5,5.5,5.8,6.5)), y = -(PCAloadings$PC2*c(4.8,4.4,4.15,5.35,5,4.7,5.75,6)+c(0,0,0,0,0,0,0,0)),
                          label = PCAloadings$Variables, color="black",fontface =2,size=4.5)

@@

all_spp(PC)

@@

breed_1 <- ggplot(breed1,aes(x= Breeding_system,y= emmean,
               label = .group)) +theme_minimal()+
  geom_point(data = breed1, aes(x = Breeding_system, y = emmean, color = Breeding_system),size   = 4) +
  geom_errorbar(aes(ymin  =  lower.CL,ymax  =  upper.CL,  color = Breeding_system),
                width =  0.25,size =  0.8)+ylab("Least square mean")+xlab("")+ theme(legend.position = "none")+ theme_ms()+
  scale_color_manual(values=c("forestgreen","orange", "black"))+ labs(title = "PC1",subtitle = "A Breeding system")+
  theme(plot.title = element_text(hjust=0.5, size=25, face="bold"))+
  geom_text(nudge_x = c(0), nudge_y = c(0.5),color   = "black")

@@

breed_2 <- ggplot(breed2,aes(x= Breeding_system,y= emmean,
             label = .group)) +theme_minimal()+
  geom_point(data = breed2, aes(x = Breeding_system, y = emmean, color = Breeding_system),size   = 4) +
  geom_errorbar(aes(ymin  =  lower.CL,ymax  =  upper.CL,  color = Breeding_system),
                width =  0.25,size =  0.8)+ylab("")+xlab("")+ theme(legend.position = "none")+ theme_ms()+
  scale_color_manual(values=c("forestgreen","orange", "black")) + labs(title = "PC2",subtitle = "B Breeding system")+
  theme(plot.title = element_text(hjust=0.5,size=25, face="bold") )+
  geom_text(nudge_x = c(0), nudge_y = c(0.35),color   = "black")

>>>>

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=10, fig.width=10, fig.cap="\\textbf{Fig. S7.} Phylogenetically informed principal component analysis (pPCA) for all plant species with the first (flower number - flower size axis) and third (style length - pollinator dependence axis) principal component that explained most trait variation."}
setwd("~/R_Projects/Reproductive Traits")

#Invert axes
PC$S[,1] <- -PC$S[,1]
PC$S[,3] <- -PC$S[,3]

pc3 <- function(PC, x="PC1", y="PC3") {
  # PC being a prcomp object
  data <- data.frame(PC$S)
  plot <- ggplot(data, aes_string(x=x, y=y)) #generate plot
  dat <- data.frame(x = data[,x], y = data[,y])

@@

  plot <- plot + xlab(paste("Flower number - flower size axis ", "(",(percentage[1]),"%",")", sep = "")) #XLAB
  plot <- plot + ylab(paste("Style length - pollinator dependence axis ", "(",(percentage[3]),"%",")", sep = "")) #YLAB

@@

  plot <- plot + theme_bw() #+ annotation_raster(tree_1, xmin = -4.1, xmax = -2.85,ymin = 2.05, ymax = 4.05)+
 #  annotation_raster(style_l_1, xmin = 2.05, xmax = 2.85,ymin = 2.35, ymax = 3.5) +
 # annotation_raster(style_s_1, xmin = -2.6, xmax = -1.9,ymin = -2.65, ymax = -1.7) +
 # annotation_raster(ovule_h_1, xmin = 3.5, xmax = 4.2,ymin = -0.45, ymax = 0.50) +
 # annotation_raster(ovule_l_1, xmin = -4.2, xmax = -3.5,ymin = -0.45, ymax = 0.5) +
 # annotation_raster(selfing_n_1, xmin = -0.75, xmax = 1.2,ymin = 1.8, ymax = 4.5) +
 # annotation_raster(selfing_h_1, xmin = -0.30, xmax = 0.15,ymin = -3.9, ymax = -3.1) +
 # annotation_raster(single_flower_1, xmin = 3.3, xmax = 4.1,ymin = -1.7, ymax = -0.7) +
 # annotation_raster(many_flowers_1, xmin = -4.20, xmax = -3.40,ymin = 0.8, ymax = 1.45) +
 # annotation_raster(large_flower_1, xmin = 3.15, xmax = 4.15,ymin = 1.3, ymax = 2.55) +
 # annotation_raster(flower_small_1, xmin = -3.50, xmax = -2.55,ymin = -2.1, ymax = -1.60) +
 # annotation_raster(herb_1, xmin = 2.6, xmax = 3,ymin = -2.8, ymax = -2)

      #ADD LABELS
  rownames(PC$L) <- c("Selfing", "Flower number", "Flower Size", "Style length", "Ovule number", "Plant height" )

  PCAloadings <- data.frame(Variables = rownames(PC$L), PC$L)
 plot <- plot + annotate("text", x = -(PCAloadings$PC1*c(4.6,4.95,5.5,4.8,7.1,5.5)), y = -(PCAloadings$PC2*c(-3.2,4.5,3.75,8.4,5,5.5)+c(0,0,0,0,-0.08,0)),
                         label = PCAloadings$Variables, color="black",fontface =2,size=4.5)

@@

pc3(PC)

@@

breed_1 <- ggplot(breed1,aes(x= Breeding_system,y= emmean,
               label = .group)) +theme_minimal()+
  geom_point(data = breed1, aes(x = Breeding_system, y = emmean, color = Breeding_system),size   = 4) +
  geom_errorbar(aes(ymin  =  lower.CL,ymax  =  upper.CL,  color = Breeding_system),
                width =  0.25,size =  0.8)+ylab("Least square mean")+xlab("")+ theme(legend.position = "none")+ theme_ms()+
  scale_color_manual(values=c("forestgreen","orange", "black"))+ labs(title = "PC1",subtitle = "A Breeding system")+
  theme(plot.title = element_text(hjust=0.5, size=26, face="bold"), plot.subtitle=element_text(size=16))+
  geom_text(nudge_x = c(0), nudge_y = c(0.5),color   = "black")

@@

breed_2 <- ggplot(breed2,aes(x= Breeding_system,y= emmean,
             label = .group)) +theme_minimal()+
  geom_point(data = breed2, aes(x = Breeding_system, y = emmean, color = Breeding_system),size   = 4) +
  geom_errorbar(aes(ymin  =  lower.CL,ymax  =  upper.CL,  color = Breeding_system),
                width =  0.25,size =  0.8)+ylab("")+xlab("")+ theme(legend.position = "none")+ theme_ms()+
  scale_color_manual(values=c("forestgreen","orange", "black")) + labs(title = "PC2",subtitle = "B Breeding system")+
  theme(plot.title = element_text(hjust=0.5,size=26, face="bold"),plot.subtitle=element_text(size=16))+
  geom_text(nudge_x = c(0), nudge_y = c(0.35),color   = "black")
